## Processing of the results of ARC analysis run
========================================================

### Script for processing the table from ARC analysis (typically the result of tree_test.pl run) and draw the linear regression

By Viacheslav Shalisko 2015

    This script forms part of ARC analysis software
    Copyright (C) 2015  Viacheslav Shalisko

    This program is free software: you can redistribute it and/or modify
    it under the terms of the GNU General Public License as published by
    the Free Software Foundation, either version 3 of the License, or
    (at your option) any later version.

    This program is distributed in the hope that it will be useful,
    but WITHOUT ANY WARRANTY; without even the implied warranty of
    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
    GNU General Public License for more details.

    You should have received a copy of the GNU General Public License
    along with this program.  If not, see <http://www.gnu.org/licenses/>.
    
-------------------

### 1) Load and check the source data

* Read source table (.csv), the table should have columns "intersect_score" for value weighted overlap of areas and "root distance" for inverse distance from root in units or branch length (the lowest) or estimated node age
* Replace "incomplete"" values with "-1" (in order to exclude cases without numeric value of weighted area intersect)
* Output table dmensions & summary table for control

```{r}
result_table <- read.csv("20150221_prueba.csv", stringsAsFactors=FALSE)

result_table[result_table$intersect_score == "incomplete",5] = "-1" 
result_table[,"intersect_score"] <- sapply(result_table[,"intersect_score"], as.numeric)

dim(result_table)
summary(result_table)
```

### 2) Select data for analysis

* Exclude lines with negative weighted area intersect values (the negative values are likely the -1 codes from previous step)
* Exclude lines with complete overlay (this is a temporal solution and should be uded with care). In case that phylogenetic tree include several nodes for the same specie, this will give readings of complete overlay, so when excluding the complete overlay scores, we exclude effect from case where several terminal nodes represent the same taxa, note that there is no problem for interspecific area comparisons, as the weighting effect compensate presence of several nodes instead of one. Sometimes the complete overlay areas could really be observed for species pair, when one species is located completely within area of distribution of another species. That´s why the analysis script require further improvement in this aspect.
* Output table dmensions & summary table for control

```{r}
useful_data <- result_table[result_table$intersect_score >= 0 & result_table$intersect_score < 1,]

dim(useful_data)
summary(useful_data)
```

### 3) Represent the node points in the scatterplot of node age (or inverse distance from the root) vs. area intersect and fit the linear regression

* Provide the points scatterplot
* Fit the linear regression and draw line
* Output the summary of regression values

```{r fig.width=10, fig.height=10}
y <- useful_data$intersect_score
x <- useful_data$root_distance

plot(x, y, type="p", col="gray10", lwd=1, pch=1,
     main="Scatterplot weighted intersect score vs. node age", 
     xlab="node age", ylab="weighted intersect score ")

linear_regression <- lm(y ~ x)
abline(linear_regression, col="red", lwd=2)

sequence <- data.frame(x=(seq(min(x), max(x), length.out = length(x))))

preds <- predict(linear_regression, newdata = sequence, 
                 interval = "confidence", level = 0.95, type="response")

lines(sequence[,1],preds[,2],col="gray",lty=2) 
lines(sequence[,1],preds[,3],col="gray",lty=2) 
```

#### Intercept value: `r linear_regression$coefficients[1]`
#### Slope value: `r linear_regression$coefficients[2]`

```{r}
summary(linear_regression)
anova(linear_regression)
```

### 4) Experiments with GLM
* Study the dependent variable distibution (normal, gamma)
* From parametric distributions Gamma seems to be better choice, but it can not be ajustid when x = 0, so small data alteration is required (0 values are substituded to 1e-10, and tha general shape of distribution as ajusted excluding 0)
* Use GLM module to fit the linear regression with data following Gamma distribution


```{r fig.width=10, fig.height=10, warning = FALSE}
library(fitdistrplus)

f_norm <- fitdist(y, "norm")
summary(f_norm)

# excluding 0 values (not a good trick, 
# but this allows to get shape in the first iteration)
yf <- y[y > 0]  
f_gamma <- fitdist(yf, "gamma")
summary(f_gamma)

# as gamma distribution cannot be fitted with x = 0, all this data 
# are changed to extremely low x = 1e-10
# the rate parameter of distribution is fixed to value from the first iteration
yf1 <- y + 1e-10
f1_gamma <- fitdist(yf1, "gamma", 
                    start=list(shape=f_gamma[["estimate"]][["shape"]]),
                    fix.arg=list(rate=f_gamma[["estimate"]][["rate"]]))
summary(f1_gamma)

hist(y, col="grey", prob=TRUE)
lines(density(y), lty="dotted", lwd=1, col="black")           # add a density estimate with defaults
lines(density(y, adjust=3), lty="dotted", lwd=2, col="red")   # add another "smoothed" density
curve(dnorm(x, mean=f_norm[["estimate"]][["mean"]], 
            sd=f_norm[["estimate"]][["sd"]]), 
            col="darkblue", lwd=2, add=TRUE, yaxt="n")
curve(dgamma(x, f_gamma[["estimate"]][["shape"]], 
             rate=f_gamma[["estimate"]][["rate"]]), 
            col="blue", lwd=2, add=TRUE, yaxt="n")
curve(dgamma(x, f1_gamma[["estimate"]][["shape"]], 
             rate=f_gamma[["estimate"]][["rate"]]), 
            col="darkgreen", lwd=2, add=TRUE, yaxt="n")

legend(0.7,2,
    c("density fit","smmothed density fit ",
      "normal","gamma (w/o 0)","gamma (with ajust of 0) "), 
    lty=c(2,2,1,1,1),
    lwd=c(1,2,2,2,2),
    col=c("black","red","darkblue","blue","darkgreen")) 

plot(x, y, type="p", col="gray10", lwd=1, pch=1,
     main="Scatterplot weighted intersect score vs. node age", 
     xlab="node age", ylab="weighted intersect score ")

g_linear_model <- glm(yf1 ~ x,
                      family=Gamma(link=identity))
abline(g_linear_model, col="blue", lwd=2)

```

#### Intercept value: `r g_linear_model$coefficients[1]`
#### Slope value: `r g_linear_model$coefficients[2]`

```{r}

summary(g_linear_model)
anova(g_linear_model)

```

### 5) Compare simple linear fit with GLM fit (Gamma)

* The AIC comparison (the lower value is better)
* The BIC comparison
* Log-Likelihood values

```{r}
# Akaike's Information Criterion
AIC(linear_regression,g_linear_model)
# Bayasian Information Criterion
BIC(linear_regression,g_linear_model)
# log-likelihood as AIC/BIC
logLik(linear_regression)
logLik(g_linear_model)
```
